---
output:
  pdf_document:
geometry: margin = 0.5in
params:
  city: input$city
  tcity: input$tCity
  region: input$region
  pop: input$pop
  budget: input$Q2_3TotTreBud
  trees: input$Q6_18PubTr
  totBud: input$Q2_1TotBud
  budmet: input$Q2_10CurrNeeds
  need: input$Q2_10Need
  admin: input$Q2_13AdminSupr
  trplan: input$Q2_13TrPlan
  trprun: input$Q2_13TrPrun
  trrem: input$Q2_13TrRem
  othr: input$Q2_13Othr
  trboard: input$Q3_2AuthTrBd
  wrtstgpln: input$Q3_9WrtStgPln
  stgplnupdt: input$Q3_9Updt
  ordn: input$Q3_5MuniOrd
  ordnUpdt: input$Q3_5YrUpdt
  trinv: input$Q6_1TrInv
  trinvUpdt: input$Q6_3YrUpd
  acp: input$Q1_3AcMMP
  acgs: input$Q1_3AcMGS
  acmpwt: input$Q1_3AcMPWTr
  havcg: input$Q6_15CanGl
  curcn: input$Q6_16CurCn
  cangl: input$Q6_16CanGl
  glyr: input$Q6_16YrToGl
  strtr: input$Q6_19StrTr
  strtrp: input$Q6_20StrTr
  prktr: input$Q6_19PrkTr
  prktrp: input$Q6_20PrkTr
  munitr: input$Q6_19MuniTr
  munitrp: input$Q6_20MuniTr
  miles: input$Q1_3MiMMS
  cipc: input$Q7_4CurCyc
  dipc: input$Q7_4DesCyc
  active: input$Q7_2SysSch
  prun: input$Q7_4AppPrun
  ner: ner
  mwr: mwr
  sr: sr
  wr: wr
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir="//uwsp.edu/files/CNR/Projects/MuniDashboard/R Code")
```

```{r, include=FALSE}
surveyData<-read.csv("TreeScoreValues.csv",header=TRUE,sep=",")
```

```{r, include=FALSE}
reg<-6
ifelse(params$region=="Midwest (Red)",reg<-1,
       ifelse(params$region==params$ner,reg<-2,
              ifelse(params$region==params$sr,reg<-3,
                     ifelse(params$region==params$wr,reg<-4,reg<-6))))
```

\pagenumbering{gobble}

# MyTreeScore Indicators `r ifelse(params$city=="Other",params$tcity,params$city)`  
\noindent\hspace*{11cm} **US Census Categories**

**Indicators & Categories** | **MyTreeScore** | **Regional** | **National** | **Unit**|
-----|:-----:|:-----:|:-----:|:-----:|
**Tree Budget & Need**| | | | |
Budget per Capita|`r ifelse(is.na(params$budget)|is.na(params$pop),'N/A',format((params$budget/params$pop),digits=2,nsmall=2))`|`r surveyData$dollarsPerCapita[reg]`|`r surveyData$dollarsPerCapita[5]`|(\$/capita)|
Budget per Public Tree|`r ifelse(is.na(params$budget)|is.na(params$trees),'N/A',format((params$budget/params$trees),digits=2,nsmall=2))`|`r surveyData$dollarsPerPublicTree[reg]`|`r surveyData$dollarsPerPublicTree[5]`|(\$/tree)|
Budget percent of Total Municipal Budget|`r ifelse(is.na(params$budget)|is.na(params$totBud),'N/A',format((params$budget/params$totBud)*100,digits=2,nsmall=2))`|`r surveyData$percentOfMuniBud[reg]`|`r surveyData$percentOfMuniBud[5]`|(%)|
Budget percent of Identified Need|`r ifelse(is_empty(params$budmet),'N/A',ifelse(params$budmet=='Yes',100,params$need))`|`r surveyData$budgetNeeds[reg]`|`r surveyData$budgetNeeds[5]`|(%)|
||||||
**Resource Budget Allocation**|||||
Administration|`r ifelse(is.na(params$admin),'N/A',params$admin)`|`r surveyData$raAdmin[reg]`|`r surveyData$raAdmin[5]`|(%)|
Planting|`r ifelse(is.na(params$trplan),'N/A',params$trplan)`|`r surveyData$raPlan[reg]`|`r surveyData$raPlan[5]`|(%)|
Pruning|`r ifelse(is.na(params$trprun),'N/A',params$trprun)`|`r surveyData$raPrun[reg]`|`r surveyData$raPrun[5]`|(%)|
Removal|`r ifelse(is.na(params$trrem),'N/A',params$trrem)`|`r surveyData$raRem[reg]`|`r surveyData$raRem[5]`|(%)|
Other|`r ifelse(is.na(params$othr),'N/A',params$othr)`|`r surveyData$raOther[reg]`|`r surveyData$raOther[5]`|(%)|
||||||
**Community Trees & Governance**|||||
Tree Board or Related Group Exists|`r ifelse(is_empty(params$trboard),'N/A',ifelse(params$trboard=='Yes',100,0))`|`r surveyData$treeBoard[reg]`|`r surveyData$treeBoard[5]`|(% yes)|
Written Strategic Plan Exists|`r ifelse(is_empty(params$wrtstgpln),'N/A',ifelse(params$wrtstgpln=='Yes',100,0))`|`r surveyData$writtenStratPlan[reg]`|`r surveyData$writtenStratPlan[5]`|(% yes)|
Written Strategic Plan up to Date|`r ifelse(is_empty(params$wrtstgpln),'N/A',ifelse(is.na(params$stgplnupdt),'N/A',ifelse((as.numeric(format(Sys.Date(), '%Y'))-params$stgplnupdt)<5,100,0)))`|`r surveyData$writtenStratPlanUpToDate[reg]`|`r surveyData$writtenStratPlanUpToDate[5]`|(% yes)|
Tree Ordinance Exists|`r ifelse(is_empty(params$ordn),'N/A',ifelse(params$ordn=='Yes',100,0))`|`r surveyData$ordinance[reg]`|`r surveyData$ordinance[5]`|(% yes)|
Tree Ordinance up to Date|`r ifelse(is_empty(params$ordn),'N/A',ifelse(is.na(params$ordnUpdt),'N/A',ifelse((as.numeric(format(Sys.Date(), '%Y'))-params$ordnUpdt)<5,100,0)))`|`r surveyData$ordinanceYear[reg]`|`r surveyData$ordinanceYear[5]`|(% yes)|
||||||
**Tree Inventory & Space  Allocation**|||||
Tree Inventory Exists|`r ifelse(is_empty(params$trinv),'N/A',ifelse(params$trinv=='Yes',100,0))`|`r surveyData$treeResourceInventoryExists[reg]`|`r surveyData$treeResourceInventoryExists[5]`|(% yes)|
Tree Inventory up to Date|`r ifelse(is_empty(params$trinv),'N/A',ifelse(is.na(params$trinvUpdt),'N/A',ifelse(as.numeric(format(Sys.Date(), '%Y'))-params$trinvUpdt<5,100,0)))`|`r surveyData$treeResourceInventory[reg]`|`r surveyData$treeResourceInventory[5]`|(% yes)|
Public Green Space Area|`r ifelse((is.na(params$acp)&is.na(params$acgs)&is.na(params$acmpwt))|is.na(params$pop),'N/A',format(((params$acp+params$acgs+params$acmpwt)*4046.86)/params$pop,digits=2,nsmall=0))`|`r surveyData$greenSpaceAreaMeters[reg]`|`r surveyData$greenSpaceAreaMeters[5]`|(sq. m / capita)|
Public Green Space Area|`r ifelse((is.na(params$acp)&is.na(params$acgs)&is.na(params$acmpwt))|is.na(params$pop),'N/A',format(((params$acp+params$acgs+params$acmpwt)*43560)/params$pop,digits=2,nsmall=0))`|`r surveyData$greenSpaceAreaFeet[reg]`|`r surveyData$greenSpaceAreaFeet[5]`|(sq. ft / capita)|
||||||
**Tree Canopy Assessment**|||||
Tree Canopy Goal Exists|`r ifelse(is_empty(params$havcg),'N/A',ifelse(params$havcg=='Yes',100,0))`|`r surveyData$canCovGoal[reg]`|`r surveyData$canCovGoal[5]`|(% yes)|
Current Tree Canopy|`r ifelse(is_empty(params$havcg),'N/A',ifelse(is.na(params$curcn),'N/A',ifelse(params$havcg=='Yes',params$curcn,'N/A')))`|`r surveyData$percentCurCan[reg]`|`r surveyData$percentCurCan[5]`|(%)|
Tree Canopy Cover Goal Achievement|`r ifelse(is_empty(params$havecg),'N/A',ifelse(is.na(params$curcn)|is.na(params$cangl),'N/A',ifelse(params$havcg=='Yes',format((params$curcn/params$cangl)*100,digits=2,nsmall=0),'N/A')))`|`r surveyData$percentCanGoal[reg]`|`r surveyData$percentCanGoal[5]`|(% Goal Met)|
Years to Achieve Tree Canopy Goal|`r ifelse(is_empty(params$havcg),'N/A',ifelse(params$havcg=='Yes',(params$glyr-as.numeric(format(Sys.Date(), '%Y'))),'N/A'))`|`r surveyData$yearsToGoal[reg]`|`r surveyData$yearsToGoal[5]`|(years)|
||||||
**Tree Stocking Level**|||||
Street Tree Stocking Level Attainment|`r ifelse(is.na(params$strtr)|(is.na(params$strtr)&is.na(params$strtrp)),'N/A',format((params$strtr/(params$strtr+params$strtrp))*100,digits=2,nsmall=0))`|`r surveyData$strTrStockingLevel[reg]`|`r surveyData$strTrStockingLevel[5]`|(% Attained)|
Park Tree Stocking Level Attainment|`r ifelse(is.na(params$prktr)|(is.na(params$prktr)|is.na(params$prktrp)),'N/A',format((params$prktr/(params$prktr+params$prktrp))*100,digits=2,nsmall=0))`|`r surveyData$prkTrStockingLevel[reg]`|`r surveyData$prkTrStockingLevel[5]`|(% Attained)|
Municipal Tree Stocking Level Attainment|`r ifelse(is.na(params$munitr)|(is.na(params$munitr)&is.na(params$munitrp)),'N/A',format((params$munitr/(params$munitr+params$munitrp))*100,digits=2,nsmall=0))`|`r surveyData$munTrStockingLevel[reg]`|`r surveyData$munTrStockingLevel[5]`|(% Attained)|
Public Street Tree Density|`r ifelse(is.na(params$strtr)|is.na(params$miles),'N/A',format((params$strtr/params$miles)*100,digits=2,nsmall=0))`|`r surveyData$pubStTrDensity[reg]`|`r surveyData$pubStTrDensity[5]`|(Trees / Mile)|
||||||
**Active (Systematic) Management**|||||
Current Inspection and Pruning Cycle|`r ifelse(is_empty(params$prun),'N/A',ifelse(is.na(params$cipc),'N/A',params$cipc))`|`r surveyData$currentPrunInsCyc[reg]`|`r surveyData$currentPrunInsCyc[5]`|(years)|
Desired Inspection and Pruning Cycle|`r ifelse(is_empty(params$prun),'N/A',ifelse(is.na(params$dipc),'N/A',params$dipc))`|`r surveyData$desiredPrunInsCyc[reg]`|`r surveyData$desiredPrunInsCyc[5]`|(years)|
Time Inspection and Pruning Off Cycle|`r ifelse(is_empty(params$prun),'N/A',ifelse(is.na(params$cipc)|is.na(params$dipc),'N/A',params$cipc-params$dipc))`|`r surveyData$yearsOfPrunInsCyc[reg]`|`r surveyData$yearsOfPrunInsCyc[5]`|(years)|
Attainment of Inspection and Pruning Cycle|`r ifelse(is_empty(params$prun),'N/A',ifelse(is.na(params$cipc)|is.na(params$dipc),'N/A',format((params$dipc/params$cipc)*100,digits=2,nsmall=0)))`|`r surveyData$percentAttPrunInsCyc[reg]`|`r surveyData$percentAttPrunInsCyc[5]`|(% Attained)|
Active Management Level of Tree Population|`r ifelse(is.na(params$active),'N/A',params$active)`|`r surveyData$activeManagement[reg]`|`r surveyData$activeManagement[5]`|(%)|

\noindent\hspace*{1cm} The analysis was calculated on `r Sys.Date()` using the MyTreeScore Dashboard &copy; 2019  
\noindent\hspace*{1cm} N/A = not applicable, community does not have this value  
\noindent\hspace*{1cm} Regional and National US Census from 2014 Municipal Tree Care and Management in the United States  
\noindent\hspace*{1cm} (Search Hauer Muni or <https://www.uwsp.edu/cnr/Pages/Forestry---MTCUS.aspx> for more information)